Detection and Classification of Supernova Gravitational Waves Signals: A Deep Learning Approach

### 2020-03-25 Response_to_PRD_reviewers.txt

We would like to express our gratitude to the reviewers for reading 
through the paper thoroughly and their positive and constructive 
comments. Below are our response to the comments.

Report of Referee A 
------------------------------------------------------------------------------
The manuscript aims at the classification of gravitational wave
signals buried in noise through a convolutional neural network that
takes time series data from gravitational wave detectors as input and
works independently. The authors took into account simulated
waveforms, generated at various distances, for which the explosion
mechanism is modelled by the neutrino-driven mechanism and the
magneto-rotational mechanism. After the training of the convolutional
neural network, the authors estimated the performance of the algorithm
by constructing the receiver operator characteristic curves. The
waveforms excluded from the training session of the convolutional
neural network are used as an extra test for the performance of the
trained CNN. The manuscript is well organised and clearly written, it
offers a complete overview of the state of art and it contains a
significant step forward for the classification of supernova
gravitational waves signals.

Response: Thank you so much for the very positive comments. 
          They are much appreciated.
------------------------------------------------------------------------------

Report of Referee B
------------------------------------------------------------------------------
The authors present a novel analysis pipeline (at the heart of which
is a convolutional neural network - CNN) for the detection and
classification of gravitational waves (GWs) from core-collapse
supernovae (CCSNe). Employing said pipeline, the authors suggest GWs
from Galactic CCSNe are likely to be detected with even the current
generation of ground-based GW detectors. If robust, this result would
provide an improvement in sensitivity by a factor of a few over
similar studies using excess-power methods, which is exciting and much
needed.

The particulars of the conditions for which the flagship paper results
are quoted, however, cause concern. A Galactic CCSN is a once in a
lifetime event, and the first "detection" of GWs from stellar collapse
would certainly not be claimed at any less than 5 sigma. As a
consequence, any results presented at FAP=1% or FAP=10% on
detectability seem a little spurious.

Going through the paper and finding the equivalent
"detection/classification distances" at FAP=0.1% (the most realistic
case presented, and a little over 3 sigma):

* [Fig. 5a] At 20kpc, TAP_HLVK ~ 1%, and TAP_H+L+VK ~ 8%

* [Fig. 5b] At 20kpc,   TAP_HLVK ~ 92% / TAP_H+L+VK ~ 95%
                At 60kpc,   TAP_HLVK ~ 55% / TAP_H+L+VK ~ 80%
                At 100kpc, TAP_HLVK ~ 16% / TAP_H+L+VK ~ 16%

* [Fig. 6a] At 10kpc, TAP_HLVK ~ 12% / TAP_H+L+VK ~ 41%

* [Fig. 6b] At 10kpc, TAP_HLVK ~ 96% / TAP_H+L+VK ~ 98%

* [Fig. 9a] At 10kpc, TAP_HLVK ~ 12% / TAP_H+L+VK ~ 41%

* [Fig. 9b] At 10kpc, TAP_HLVK ~ 96% / TAP_H+L+VK ~ 97%

These numbers (and the above commentary) suggest that a different
region of parameter space might be more appropriate to study here.
Assuming TAP>50% and FAP<0.3% as a (generous) lower limit on realistic
detection claims, it would be instructive to instead show:

* [Fig. 5a] TAP_HLVK vs FAP_HLVK and TAP_H+L+VK vs FAP_H+L+VK between
0.001% < FAP < 1% at 0.2kpc (Betelgeuse), 2kpc, and 10kpc (typical
"Galactic SN" distance)

* [Fig. 5b] TAP_HLVK vs FAP_HLVK and TAP_H+L+VK vs FAP_H+L+VK between
0.001% < FAP < 1% at 2kpc, 20kpc, 60kpc

* [Fig. 6a] TAP_HLVK and TAP_H+L+VK as a function of distance between
0.5kpc - 20kpc at FAP=0.1%, 0.01%, and 0.001%

* [Fig. 6b] TAP_HLVK and TAP_H+L+VK as a function of distance between
0.5kpc - 200kpc at FAP=0.1%, 0.01%, and 0.001%

* [Fig. 9a] TAP_HLVK and TAP_H+L+VK as a function of distance between
0.5kpc - 20kpc at FAP=0.1%, 0.01%, and 0.001%

* [Fig. 9b] TAP_HLVK and TAP_H+L+VK as a function of distance between
0.5kpc - 200kpc at FAP=0.1%, 0.01%, and 0.001%

Looking at Fig. 7, however, I question if it is possible to achieve
such low FAPs with this pipeline, particularly for the neutrino
mechanism waveforms. It would be useful to remake this figure at
distances for which FAP << 0.3% is possible. If not possible, it may
be beneficial to remove the figure. If not removed, the caption and
text should be updated to note the large FAPs used.

Response: We have tested again the trained CNN with 1.2^5 test samples 
          for each distance so that we can achieved a FAP of 1%, 0.1% 
          and 0.01%.

          We appreciate that 0.001% is also one of the requested values
          of FAP and running the testing at closer distance for neutrino
          -driven mechanism is advised. 
          But we also noted that such a low value of FAP would 
          require an even larger set of test samples and a much longer 
          amount of time to generate, which maybe too computationally 
          expensive for us to do. Rerunning the simulations at closer
          distances will even be more computationally expensive.

          The changes in response to this comment includes: 
          1. Only TAP values at FAP equal to 1% are qouted in the      
             abstract and the conclusion. No detection claim is made for 
             FAP higher than 1%. 
             We understand that detection claims are advised to be made 
             at least FAP = 0.1%, but we argue that detections of CCSNe 
             signals will be made simultaneously with electromagnetic 
             observations and/or neutrino observations, and therefore 
             such low FAP may not be absolutely necessary.

          2. The plots and text in the section of results 
             (section IV and section V).

          3. We retracted claims regarding the ability of the 
             network HLVK in detecting Galactic events where the explosion 
             mechanism is neutrino-driven mechanism.

          4. number of training/test samples in the last paragraph of 
             section III has been updated to be consistent with 
             this change.

          5. We would also like to highlight the fact that the plots
             previously submitted as Fig. 9a and Fig. 9b were actually 
             duplicates of previous Fig. 6a and Fig. 6b by accident. 
             Although this mistake was naturaly corrected when we updated
             the results, we would still like to bring this mistake to the
             reviewers' attention.

Overall, the study is a novel and interesting work. It is crucial,
however, to put the authors' pipeline through more rigorous testing
(i.e. at realistic FAPs, perhaps using realistic noise) before stating
its ability to detect GWs from Galactic CCSNe. I would suggest framing
instead the paper as a follow-up pipeline, but take note that the
authors explicitly distinguish their pipeline from that shown in a
similar work by Astone+2018 as designed to take time series directly
from the detectors and find signals, rather than analyzing triggers
found by some external detection pipeline. Given this, it seems
appropriate for appreciable corrections to the text and results to be
requested.

Response: We appreciate this comment and admit that testing/training 
          the CNN using realistic noise would indeed show a more 
          realistic results.
          However, such extension of the current work would require 
          a whole new set of treatment that is beyond the current scope
          of this work. 
          We therefore updated our results with more realistic 
          values of FAP and leave tests using realistic noise for future 
          work.

In the below, I list specific comments by section, denoting the
relevant paragraph number by p.

I: Introduction

p1: LIGO, Virgo, and KAGRA are not introduced or cited here. Suggest
to move discussion of second generation detectors (and references)
from start of p2 here. Aside - has KAGRA now begun operations? If so,
it may be worth updating the text here.

Response: We have updated the text accordingly.

p2: Specifically, the sentence "The nuclear equation of state will
then stiffens by the strong nuclear force above the nuclear density
and stops the core collapse." doesn't really make sense. The equation
of state stiffens (and subsequently stops the core collapsing) as the
core density exceeds nuclear density because the strong nuclear force
becomes repulsive at very small inter-nucleon distances.

Response: Thanks for pointing this out. We have rewritten the sentense 
          to read "The nuclear equation of state will then stiffens by 
          the strong nuclear force as the core exceeds nuclear density
          and stops the core collapse."

II: Convolutional Neural Network & Fig. 1 The technical description of
the CNN pipeline is thorough and instructive - nice work!

Response: Thank you so much.

Tab. 1 caption: typo at end of first sentence, detectr --> detector

Response: This has been corrected.

III: Data

p3: For the 3D simulations, what is the observer angle for the h+,hx
used? As the GW signal for CCSNe varies with orientation of the source
with respect to the observer, it may be useful to add a distribution
of observer angles for each waveform model to the training set, so as
to account for this variance. This could also help remedy the issue
noted in p6 of small sample size of neutrino mechanism waveforms
compared to magnetorotational mechanism waveforms.

Response: the observer angles for all the studies of 3D simulations have
          been added to the text. We appreciate that our data could be 
          augmented by varying the observer angles in the data set. 
          However, we noted that while some of the used studies published 
          the information or parameters to generate waveforms at random 
          observer angles, the same information for other studies is not
          readily available. We therefore decided to use only the published
          waveforms.

Fig. 2 & Fig. 3: What are the observer angles for these waveforms? As
an aside, it could be useful to combine these plots into a 4 panel
figure.

Response: the observer angles for these figures have been added in 
          the caption.

Note: Appreciated the technical detail included on signal injection
methodology - very refreshing to see, and crucial for reproducibility.

Response: Thank you so much.

IV: Results and Discussion

I suggest the results are rerun using the parameters highlighted
above, and the text subsequently rewritten to reflect the updated
results. At the very least, statements on detecting GWs from a
Galactic CCSN should be removed (unless results with FAP < 0.3% show
this is the case), and a clear note/disclaimer on the high FAPs used
must be added.

Note: Figs. 5 and 6 are clear and well presented - very instructive.

Response: As stated above, we have updated section IV and section V.

V: Waveforms Excluded From Training

As for the results section, I suggest reruns using parameters
highlighted above, and the text subsequently rewritten to reflect the
updated results. Similarly to before, at the very least a disclaimer
on the FAPs must be added, and detection statements for FAP>0.3%
removed.

Response: As stated above, we tested again the trained CNN with 
          1.2 * 10^5 test samples to achieve FAP values of 1%, 0.1%, 
          0.01% as was done for section IV. The text and plots 
          are also updated to reflect this change. 
          Claims regarding the ability of the network HLVK in detecting 
          Galatic CCSN signals if the explosion mechanism is the 
          neutrino driven mechanism.

Fig. 8: As for Figs. 2 and 3, what are the observer angles for these
waveforms. As it is noted that h+ >> hx in the caption, I suspect the
waveform is for a polar observer, but this should be confirmed and
noted in the text.

Response: The waveforms shown in Fig. 8a and Fig. 8b are indeed for a 
          polar observer. This as well as that for Fig. 8c and Fig. 8d
          have been added in the caption.

VI: Conclusion

As previously noted, only results at FAP=10% are mentioned. As it is
not realistic to claim detection at such high FAP, the distances
presented are meaningless. This should be rewritten, pointing instead
to the results at FAP=0.1%. If Galactic/Magellanic Cloud detections
are no longer feasible, the abstract (and other relevant sections)
should be rewritten to reflect that.

Response: We updated the text to be consistent with the uses of more realistic 
          FAP and retracted claims of detection of Galactic event with HLVK.
------------------------------------------------------------------------------

### 2020-04-17 main.tex
comments incorporated, samples increased to 1.08e6

\begin{abstract}
We demonstrate the application of a convolutional neural network to the
gravitational wave signals from core collapse supernovae. Using simulated time
series of gravitational wave detectors, we show that based on the explosion
mechanisms, a convolutional neural network can be used to detect and classify
the gravitational wave signals buried in noise. For the waveforms used in the
training of the convolutional neural network, our results suggest that a
network of advanced LIGO, advanced VIRGO and KAGRA, or a network of LIGO A+,
advanced VIRGO and KAGRA is likely to detect a magnetorotational core collapse
supernovae out to the Large or even Small Magellanic Clouds. A Galactic event 
is likely to be detectable with a network of H+L+VK if the explosion mechanism 
is the neutrino-driven mechanism.  By testing the convolutional neural network 
with waveforms not used for training, we show that the true alarm probabilities 
are $50\%$ and $73\%$ at $50$ kpc for waveforms $\text{R3E1AC}$ and 
$\text{R4E1FC\_L}$ for networks of HLVK and H+L+VK respectively.  
For waveforms $\text{s}20$ and $\text{SFHx}$ at $10$ kpc, 
the true alarm probabilities are $64\%$ for a network of H+L+VK. 
All at false alarm probability equal to $1\%$. 
\end{abstract}

@@

%
% plot of example magneto waveforms
%
\begin{figure*}
     \begin{center}
        \subfigure[]{
            \label{fig:magsample1}
            \includegraphics[width=0.315\textwidth]{abd_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:magsample2}
            \includegraphics[width=0.315\textwidth]{Dim_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:magsample3}
            \includegraphics[width=0.315\textwidth]{richer_wave.png}%
        }
    \end{center}
    \caption{Examples of simulated waveforms for which the explosion is modelled by
the magnetorotational mechanism. The left panel shows a waveform 
from~\cite{abdikamalov2014measuring} for an equatorial observer. The progenitor is $12\text{M}_\odot$ with
differential rotation parameter $A=10^4~\text{km}$. The initial angular velocity
at the core $\Omega_c$ is $5.0~\text{rads/s}$. The middle panel shows a waveform
from~\cite{dimmelmeier2008gravitational}. The simulation corresponds to a
progenitor of $15\text{M}_\odot$ with $A=10^5$ km and $\Omega_c =
4.56~\text{rads/s}$.  The right panel shows a waveform
from~\cite{richers2017equation} where the progenitor is $12\text{M}_\odot$, $A$
and $\Omega_c$ are $10^4$ km and $3.0~\text{rad/s}$ respectively. In all panels,
only the $h_{+}$ polarisations are shown because the simulations are
axis-symmetric or 2D, and therefore described by only one polarisation.  The
sources are assumed to be at a distance of $10$ kpc. The x-axes show the time after 
core bounce. The waveforms shown in all panels are for an observer in the equatorial plane.\label{fig:magwaveforms}} 
\end{figure*}
%
% plot of example neutrino waveforms
%
\begin{figure*}
     \begin{center}
        \subfigure[]{
            \label{fig:neusample1}
            \includegraphics[width=0.315\textwidth]{and_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:neusample2}
            \includegraphics[width=0.315\textwidth]{Radice_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:neusample3}
            \includegraphics[width=0.315\textwidth]{Yak_wave.png}%
        }
    \end{center}
    \caption{Examples of simulated waveforms for which the explosion mechanism is
modelled by the neutrino-driven mechanism. From the left to the right, the
panels show waveforms from~\cite{10.1093mnrasstz990, 
radice2019characterizing, yakunin2017gravitational}, with progenitors of masses equal to
$15\text{M}_\odot$, $15\text{M}_\odot$ and $60\text{M}_\odot$ respectively.
For the left and middle panels, the observer is assumed to be along the pole, 
while an observer along the equator is assumed for the right panel.
The sources are assumed to be at $10$ kpc from earth. The x-axes show the time after core bounce.
The waveform shown in the left panel is for a polar observer, while the other panels show that 
for an equatorial observer.
\label{fig:neuwaveforms}} 
\end{figure*}
%
% discuss the reason of using multiple waveform families
%
Introducing waveforms from a variety of studies introduces a distribution of
the waveforms that covers a larger parameter space, making the data samples
harder for the \ac{CNN} to learn. As a result, the performance of the trained
\ac{CNN} on the testing samples may appear to be worse than that if the data
were generated using only a subset of the waveforms. However, a larger
parameter space provides the advantage of forcing the \ac{CNN} to learn the
features common among the waveforms rather than simply a subset of the
waveforms.  The trained network can thus perform better when presented with
waveforms with unexpected features, which is more likely in reality.

@@

%
% data augmentation
%
As a general rule, the performance of a \ac{CNN} can always be enhanced by
increasing the volume of training data. Therefore we augment our data set by
iterating over $\bold{S}(t)$ until for each waveform in $\bold{S}(t)$, we have
$31$ data samples $\bold{d}$ generated using the procedure described above.
This step is essential for the \ac{CNN} to learn the features of the waveforms
and to identify them effectively under different noise scenarios and starting
times.  After that, we generate independent noise realisations for the
background noise class. The entire data set $\bold{D}$ has
roughly $l = 3.0 \times 10^{5}$ data samples, where each of the 3 classes
contains approximately $1.0\times10^{5}$ data samples.
%
%\begin{equation}\label{eq:finaldata}
%\bold{D}(\boldsymbol{\alpha}, \boldsymbol{\delta}, \boldsymbol{\delta_t}, \bold{t}, \bold{x}) =  \left(\begin{array}{c}
%\bold{d}_1(\alpha, \delta, \delta_t, t, x) \\
%\vdots \\
%\bold{d}_o(\alpha, \delta, \delta_t, t, x) \\
%\vdots \\
%\bold{d}_l(t) \\
%\end{array}\right).
%\end{equation}
%
%In the above equation, the subscript $o$ is the sum of the data samples in
%which there is a \ac{GW} signal in the time series.~\cm{this notation is very
%confusing and not really required. Why do you have to stack the waveforms up
%like this? Having to define things like $o$ and $l$ seems messy.} The final
%step is to whiten $\bold{D}$ using the power spectral densities. 
%
% the general training procedure
%
In Fig.~\ref{fig:sample}, we show a representative example of a pre-processed
input time-series for a single detector. A data sample consists of such a
time-series containing a signal from the same source or simply background noise
from all the detectors in the network. The data samples are then split into 3
groups, with $1.0\times10^{4}$ samples being randomly selected for validation.
Of the remainder, $\sim 1.7\times10^{5}$ for training and $\sim 1.2\times10^{5}$ for tesing. 
The large number of testing samples is chosen so that low values of \ac{FAP} can be obtained 
(see section \ref{sec:result} and section \ref{sec:unseen}).
When the training is finished for a dataset generated at a given distance, the above described
procedure will be repeated for another distance with a different \ac{CNN} of 
the same architecture until the training for all
distances have been carried out for a \ac{GW} detector network. The entire
procedure is then repeated for another \ac{GW} detector network. 
As mentioned in the introduction, the \ac{CNN} will be trained for the two networks of
\ac{GW} detectors presented in Table~\ref{table:network}. For the remaining of
the paper, we will use their acronyms to refer to the networks, i.e., HLVK and H+L+VK.

%
% plot of representative data sample
%
\begin{figure}
\includegraphics[width=0.475\textwidth]{datasample.png}
\caption{A representative example of the simulated time-series used to
train/validate/test the \ac{CNN}. The blue shows a whitened time-series with a
signal added to Gaussian noise generated using the power spectral density of \ac{aLIGO}. 
The red curve shows the same whitened signal free of
noise. The original waveform is shown in Fig. \ref{fig:neusample1}.
The source is at a distance of $10$ kpc from earth.
\label{fig:sample}} 
\end{figure}

@@

\begin{figure*}
     \begin{center}
        \subfigure[]{
            \label{fig:ROClog1}
            \includegraphics[width=0.475\textwidth]{ROC_curves_neutrino.png}%
        }\quad
        \subfigure[]{
            \label{fig:ROClog2}
            \includegraphics[width=0.475\textwidth]{ROC_curves_mag.png}%
        }
    \end{center}
    \caption{\ac{ROC} curves showing the classification performance of the
\ac{CNN} for \ac{CCSN} signals of different explosion mechanisms at three
distances, $20$, $60$, and $100$ kpc. The left panel shows the results for the
neutrino-driven mechanism, while the panel on the right shows those for the
magnetorotational mechanism. In both panels, the solid lines are for the
network H+L+VK, while the dashed lines are for HLVK. The black diagonal lines in both plots indicate 
the worst scenario where a model has zero ability in distinguishing its input.
\label{fig:ROClog}} 
\end{figure*}

%
% discuss the accuracy as a function of distance (Fig 6)
%
We also show the classification efficiency of the \ac{CNN}  as a function of
distance.  This is done by fixing the \ac{FAP} and plotting the \ac{TAP}.  The
results for three chosen \acp{FAP} values are shown in Fig.~\ref{fig:eff}.  In
this figure, a similar trend is seen that magnetorotational signals are easier
for the \ac{CNN} to identify than neutrino-driven signals at a given distance.
For magnetorotational signals from sources located at $50$ kpc and a \ac{FAP}
of $1\%$, the \ac{CNN} achieves a \ac{TAP} of $87\%$ and $72\%$ for the
networks H+L+VK and HLVK respectively. At more restrict \acp{FAP} such as
$0.01\%$, the \ac{CNN} still achieves \acp{TAP} of $75\%$ and $62\%$ for sources at
the same distance for the two networks respectively.  For sources that are
located at a slightly further distance of $60$ kpc, the networks achieve a
\ac{TAP} of $82\%$ (H+L+VK) or $60\%$ (HLVK) at a \ac{FAP} of $1\%$. Such a range
includes distances associated with the Large and Small Magellanic Clouds and
covers the satellite galaxies in between~\cite{karachentsev2004catalog,
belokurov2007cats}.  For sources at $100$ kpc, the \acp{TAP} are no less than $50\%$ 
for H+L+VK for all chosen \acp{FAP} values, and is $61\%$ if
the \ac{FAP} is $1\%$. These numbers indicate that with
such a \ac{GW} network, it is possible to detect magnetorotational \ac{CCSN}
signals out to such a distance. For sources at $150$ and $200$ kpc, the \acp{TAP}
drop to $39\%$ and $21\%$ respectively for \ac{FAP} equal to $1\%$.
On the other hand, it is more difficult for the
\ac{CNN} to detect and classify the neutrino-driven signals, due to their
weaker amplitudes. For example, for sources at $10$ kpc, the \ac{CNN} achieves
a \ac{TAP} of $53\%$ and $22\%$ for H+L+VK and HLVK respectively if the
\ac{FAP} is $1\%$.  This means that for a \ac{GW} from a Galactic \ac{CCSN}
signal it is possible to detect and classify it with a network of H+L+VK.

@@

In practice, since the output of a \ac{CNN} is probabilities indicating how
likely the input belongs to each of the classes, it may be desired to set a
threshold probability on which the decision whether the input belong to a class is made.
For example, an input will be assigned into a certain class if the
corresponding probability is larger than the pre-selected threshold probability. In such a
scenario, the \ac{FAP} and \ac{TAP} would be affected by the choice of the
threshold. We show such a result in Fig.~\ref{fig:ROCfixed}, where we employed
a threshold probability of $95\%$. 
This value is chosen so that low values of \ac{FAP} can be maintained throughout all tested distances
as illustrated by the right y-axes of the figures.
%because for lower values, 
%there could be more than a class with predicted probabilities from the \ac{CNN} 
%larger than the threshold, while higher values could potentially rule out inputs 
%that may otherwise be correctly identified.
For H+L+VK and magnetorotational signals at $10$ kpc, the \ac{TAP} is $97\%$. If the distance
is extended to $100$ kpc, the \ac{TAP} is still larger than $50\%$. 
For HLVK, the \ac{TAP} is $95\%$ for sources at $10$ kpc and drops to $54\%$ at
$60$ kpc.
%The \acp{TAP} decrease to $10\%$ and $9\%$ for sources at $150$ and $200$ kpc.
For all distances, the \ac{CNN} maintains a \ac{FAP} no larger than $0.2\%$ for
both networks regarding magnetorotational waveforms. For neutrino-driven
signals, we show that H+L+VK has a \acp{TAP} of $39\%$ at $10$ kpc while it
is $6\%$ at $10$ kpc for HLVK with \acp{FAP} $\sim 0.1\%$. It should be noted that the results presented in this section are
averaged over all the data samples in the testing samples. The performance on
any individual waveform may vary depending on the morphology and the amplitudes
of the waveform.

>>>>

\begin{abstract}
We demonstrate the application of a convolutional neural network to the
gravitational wave signals from core collapse supernovae. Using simulated time
series of gravitational wave detectors, we show that based on the explosion
mechanisms, a convolutional neural network can be used to detect and classify
the gravitational wave signals buried in noise. For the waveforms used in the
training of the convolutional neural network, our results suggest that a
network of advanced LIGO, advanced VIRGO and KAGRA, or a network of LIGO A+,
advanced VIRGO and KAGRA is likely to detect a magnetorotational core collapse
supernovae out to the Large or even Small Magellanic Clouds. 
%A Galactic event is likely to be detectable with a network of H+L+VK 
%if the explosion mechanism is the neutrino-driven mechanism.  
By testing the convolutional neural network 
with waveforms not used for training, we show that the true alarm probability
is $66\%$ at $60$ kpc for waveforms $\text{R3E1AC}$ and 
$\text{R4E1FC\_L}$ for networks of H+L+VK.  
For waveforms $\text{s}20$ and $\text{SFHx}$ at $10$ kpc, 
the true alarm probabilities are $49\%$ and $20\%$ for a network of H+L+VK 
and HLVK respectively. 
All at false alarm probability equal to $0.1\%$. 
\end{abstract}

@@

%
% plot of example magneto waveforms
%
\begin{figure*}
     \begin{center}
        \subfigure[]{
            \label{fig:magsample1}
            \includegraphics[width=0.315\textwidth]{abd_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:magsample2}
            \includegraphics[width=0.315\textwidth]{Dim_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:magsample3}
            \includegraphics[width=0.315\textwidth]{richer_wave.png}%
        }
    \end{center}
    \caption{Examples of simulated waveforms for which the explosion is modelled by
the magnetorotational mechanism. The left panel shows a waveform 
from~\cite{abdikamalov2014measuring}. The progenitor is $12\text{M}_\odot$ with
differential rotation parameter $A=10^4~\text{km}$. The initial angular velocity
at the core $\Omega_c$ is $5.0~\text{rads/s}$. The middle panel shows a waveform
from~\cite{dimmelmeier2008gravitational}. The simulation corresponds to a
progenitor of $15\text{M}_\odot$ with $A=10^5$ km and $\Omega_c =
4.56~\text{rads/s}$.  The right panel shows a waveform
from~\cite{richers2017equation} where the progenitor is $12\text{M}_\odot$, $A$
and $\Omega_c$ are $10^4$ km and $3.0~\text{rad/s}$ respectively. In all panels,
only the $h_{+}$ polarisations are shown because the simulations are
axis-symmetric or 2D, and therefore described by only one polarisation. 
The waveforms shown in all panels are for an observer in the equatorial plane.
The sources are assumed to be at a distance of $10$ kpc. The x-axes show the time after 
core bounce.\label{fig:magwaveforms}} 
\end{figure*}
%
% plot of example neutrino waveforms
%
\begin{figure*}
     \begin{center}
        \subfigure[]{
            \label{fig:neusample1}
            \includegraphics[width=0.315\textwidth]{and_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:neusample2}
            \includegraphics[width=0.315\textwidth]{Radice_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:neusample3}
            \includegraphics[width=0.315\textwidth]{Yak_wave.png}%
        }
    \end{center}
    \caption{Examples of simulated waveforms for which the explosion mechanism is
modelled by the neutrino-driven mechanism. From the left to the right, the
panels show waveforms from~\cite{10.1093mnrasstz990, 
radice2019characterizing, yakunin2017gravitational}, with progenitors of masses equal to
$15\text{M}_\odot$, $15\text{M}_\odot$ and $60\text{M}_\odot$ respectively.
The waveform shown in the left panel is for a polar observer, while the waveforms in the other panels are 
for an equatorial observer.
The sources are assumed to be at $10$ kpc from earth. The x-axes show the time after core bounce.
\label{fig:neuwaveforms}} 
\end{figure*}
%
% discuss the reason of using multiple waveform families
%
Introducing waveforms from a variety of studies introduces a distribution of
the waveforms that covers a larger parameter space, making the data samples
harder for the \ac{CNN} to learn. As a result, the performance of the trained
\ac{CNN} on the testing samples may appear to be worse than that if the data
were generated using only a subset of the waveforms. However, a larger
parameter space provides the advantage of forcing the \ac{CNN} to learn the
features common among the waveforms rather than simply a subset of the
waveforms.  The trained network can thus perform better when presented with
waveforms with unexpected features, which is more likely in reality.

@@

%
% data augmentation
%
As a general rule, the performance of a \ac{CNN} can always be enhanced by
increasing the volume of training data. Therefore we augment our data set by
iterating over $\bold{S}(t)$ until for each waveform in $\bold{S}(t)$, we have
$31$ data samples $\bold{d}$ generated using the procedure described above.
This step is essential for the \ac{CNN} to learn the features of the waveforms
and to identify them effectively under different noise scenarios and starting
times.  After that, we generate independent noise realisations for the
background noise class. The entire data set $\bold{D}$ has
roughly $l = 1.26 \times 10^{6}$ data samples, where each of the 3 classes
contains approximately $4.2\times10^{5}$ data samples.
%
%\begin{equation}\label{eq:finaldata}
%\bold{D}(\boldsymbol{\alpha}, \boldsymbol{\delta}, \boldsymbol{\delta_t}, \bold{t}, \bold{x}) =  \left(\begin{array}{c}
%\bold{d}_1(\alpha, \delta, \delta_t, t, x) \\
%\vdots \\
%\bold{d}_o(\alpha, \delta, \delta_t, t, x) \\
%\vdots \\
%\bold{d}_l(t) \\
%\end{array}\right).
%\end{equation}
%
%In the above equation, the subscript $o$ is the sum of the data samples in
%which there is a \ac{GW} signal in the time series.~\cm{this notation is very
%confusing and not really required. Why do you have to stack the waveforms up
%like this? Having to define things like $o$ and $l$ seems messy.} The final
%step is to whiten $\bold{D}$ using the power spectral densities. 
%
% the general training procedure
%
In Fig.~\ref{fig:sample}, we show a representative example of a pre-processed
input time-series for a single detector. A data sample consists of such a
time-series containing a signal from the same source or simply background noise
from all the detectors in the network. The data samples are then split into 3
groups, with $1.0\times10^{4}$ samples being randomly selected for validation.
Of the remainder, $1.70\times10^{5}$ for training and $1.08\times10^{6}$ for tesing. 
The large number of testing samples is chosen so that low values of \ac{FAP} can be obtained 
(see section \ref{sec:result} and section \ref{sec:unseen}).
When the training is finished for a dataset generated at a given distance, the above described
procedure will be repeated for another distance with a different \ac{CNN} of 
the same architecture until the training for all
distances have been carried out for a \ac{GW} detector network. The entire
procedure is then repeated for another \ac{GW} detector network. 
As mentioned in the introduction, the \ac{CNN} will be trained for the two networks of
\ac{GW} detectors presented in Table~\ref{table:network}. For the remaining of
the paper, we will use their acronyms to refer to the networks, i.e., HLVK and H+L+VK.

%
% plot of representative data sample
%
\begin{figure}
\includegraphics[width=0.475\textwidth]{datasample.png}
\caption{A representative example of the simulated time-series used to
train/validate/test the \ac{CNN}. The blue shows a whitened time-series with a
signal added to Gaussian noise generated using the power spectral density of \ac{aLIGO}. 
The red curve shows the same whitened signal free of
noise. The optimal signal to noise ratio for this signal is $5.5$.
The original waveform is shown in Fig. \ref{fig:neusample1}.
The source is at a distance of $10$ kpc from earth.
\label{fig:sample}} 
\end{figure}

@@

\begin{figure*}
     \begin{center}
        \subfigure[]{
            \label{fig:ROClog1}
            \includegraphics[width=0.475\textwidth]{ROC_curves_neutrino.png}%
        }\quad
        \subfigure[]{
            \label{fig:ROClog2}
            \includegraphics[width=0.475\textwidth]{ROC_curves_mag.png}%
        }
    \end{center}
    \caption{\ac{ROC} curves showing the classification performance of the
\ac{CNN} for \ac{CCSN} signals of different explosion mechanisms at three
distances, $20$, $60$, and $100$ kpc. The left panel shows the results for the
neutrino-driven mechanism, while the panel on the right shows those for the
magnetorotational mechanism. In both panels, the solid lines are for the
network H+L+VK, while the dashed lines are for HLVK. The black lines in both plots indicate 
the worst scenario where a model has zero ability in distinguishing its input.
\label{fig:ROClog}} 
\end{figure*}

%
% discuss the accuracy as a function of distance (Fig 6)
%
We also show the classification efficiency of the \ac{CNN}  as a function of
distance.  This is done by fixing the \ac{FAP} and plotting the \ac{TAP}.  The
results for three chosen \acp{FAP} values are shown in Fig.~\ref{fig:eff}.  In
this figure, a similar trend is seen that magnetorotational signals are easier
for the \ac{CNN} to identify than neutrino-driven signals at a given distance.
For magnetorotational signals from sources located at $50$ kpc and a \ac{FAP}
of $0.1\%$, the \ac{CNN} achieves a \ac{TAP} of $82\%$ and $66\%$ for the
networks H+L+VK and HLVK respectively. At more restrict \acp{FAP} such as
$0.001\%$, the \ac{CNN} still achieves \acp{TAP} of $72\%$ and $55\%$ for sources at
the same distance for the two networks respectively.  For sources that are
located at a slightly further distance of $60$ kpc, the networks achieve a
\ac{TAP} of $77\%$ (H+L+VK) or $54\%$ (HLVK) at a \ac{FAP} of $0.1\%$. Such a range
includes distances associated with the Large and Small Magellanic Clouds and
covers the satellite galaxies in between~\cite{karachentsev2004catalog,
belokurov2007cats}.  For sources at $80$ kpc, the \acp{TAP} are no less than $50\%$ 
for H+L+VK for all chosen \acp{FAP} values, and is $66\%$ if
the \ac{FAP} is $0.1\%$. These numbers indicate that with
such a \ac{GW} network, it is possible to detect magnetorotational \ac{CCSN}
signals out to such a distance. For sources at $100$, $150$ and $200$ kpc, the \acp{TAP}
drop to $55\%$, $33\%$ and $16\%$ respectively for \ac{FAP} equal to $0.1\%$.
On the other hand, it is more difficult for the
\ac{CNN} to detect and classify the neutrino-driven signals, due to their
weaker amplitudes. For example, for sources at $10$ kpc, the \ac{CNN} achieves
a \ac{TAP} of $38\%$ and $10\%$ for H+L+VK and HLVK respectively if the
\ac{FAP} is $0.1\%$. In Fig.~\ref{fig:eff}, only the values of \ac{TAP} that
are larger than zero will be plotted.

@@

In practice, since the output of a \ac{CNN} is probabilities indicating how
likely the input belongs to each of the classes, it may be desired to set a
threshold probability on which the decision whether the input belong to a class is made.
For example, an input will be assigned into a certain class if the
corresponding probability is larger than the pre-selected threshold probability. In such a
scenario, the \ac{FAP} and \ac{TAP} would be affected by the choice of the
threshold. We show such a result in Fig.~\ref{fig:ROCfixed}, where we employed
a threshold probability of $90\%$. 
This value is chosen so that low values of \ac{FAP} can be maintained throughout all tested distances
as illustrated by the right y-axes of the figures.
%because for lower values, 
%there could be more than a class with predicted probabilities from the \ac{CNN} 
%larger than the threshold, while higher values could potentially rule out inputs 
%that may otherwise be correctly identified.
For H+L+VK and magnetorotational signals at $10$ kpc, the \ac{TAP} is $98\%$. 
If the distance is extended to $100$ kpc, the \ac{TAP} is still larger than $50\%$. 
For HLVK, the \ac{TAP} is $96\%$ for sources at $10$ kpc and drops to $56\%$ at
$60$ kpc.
%The \acp{TAP} decrease to $10\%$ and $9\%$ for sources at $150$ and $200$ kpc.
For all distances, the \ac{CNN} maintains a \ac{FAP} no larger than $0.3\%$ for
both networks regarding magnetorotational waveforms. For neutrino-driven
signals, we show that H+L+VK has a \acp{TAP} of $44\%$ at $10$ kpc while it
is $11\%$ at $10$ kpc for HLVK with \acp{FAP} $< 0.3\%$.
As we have chosen a large value of threshold, the values of \ac{TAP} decreases rapidly 
as the distance increases. Again, only the values of \ac{TAP} larger than zero will be plotted.
